---
title: "ORA and GSEA with Ulisse"
date: "`r Sys.Date()`"
package: Ulisse
bibliography: references.bib
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{ORA and GSEA with Ulisse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning=FALSE}
devtools::load_all()
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error    = FALSE,
  warning  = FALSE,
  #eval     = FALSE,
  message  = FALSE#,
  #fig.width = 7,
  #fig.height = 7
)
```

In this vignette we will show how to use Ulisse to perform Over-Representatin Analysis (ORA) and Gene Set Enrichment Analysis (GSEA).

# ORA

ORA is performed by means of the function `ora()`, which requires a list of one or more gene vectors (`wb`), the reference universe and a list of gene sets to test (`gsl`):

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
ora_res <- ora(wb = wb, universe = unverse, gsl = gsl)
```

# GSEA

GSEA is performed by means of the function `gsea()`, which requires a matrix with one or more columns of gene-related statistics (`rl`) and a list of gene sets to test (`gsl`):

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
gsea_res <- gsea(rl=rl, gsl = gsl)
```

By default, each column of `rl` is sorted in decreasing order.

# Visualization

The functions `plot_ora_heatmap()` and `plot_gsea_heatmap()` draw heatmaps with results of ORA and GSEA:

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
plot_ora_heatmap(ora_res, cluster_columns = F, cluster_rows = F, a = .5, na_col = "gold")
plot_gsea_heatmap(gsea_res = gsea_res, cluster_columns = F, cluster_rows = F, a = .5, na_col = "gold")
```

# Inter-operability

We provide the functions `ora2enrich()` and `gsea2enrich()` to obtain instances of "enrichResult" class of the package clusterProfiler:

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
ora_enrich <- ora2enrich(ora_res, gsl = gsl, wb = wb, universe = unverse)
gsea_enrich <- gsea2enrich(gsea_res, rl = rl, gsl = gsl)
```
